---
title: "Tree growth, mortality, and climatic stress in west-coast states, USA"
author:
  - name: "Jeremiah D. Groom"
    affiliations:
      - name: "Groom Analytics LLC"
        address: "1975 SE Crystal Lake Dr., Unit 173"
        city: "Corvallis"
        state: "OR"
        postal-code: "97333"
    orcid: "0000-0002-0110-1036"
    email: "jeremy@groomanalytics.com"
  - name: "Bryce Frank"
    affiliations:
      - name: "USDA Forest Service"
        department: "Pacific Northwest Field Station"
        address: "3625 93rd Ave SW"
        city: "Olympia"
        state: "WA"
        postal-code: "98512"
abstract: |
  Your abstract goes here. This should be a concise summary of your research question, methods, key findings, and implications. Keep it within journal word limits (typically 150-300 words).
keywords: [keyword1, keyword2, keyword3, keyword4]
date: today
format:
  docx:
    reference-doc: template_use2.docx 
    fig-width: 6
    fig-height: 4
    fig-dpi: 300
    toc: false
    number-sections: true
    highlight-style: github
execute:
  echo: false
  warning: false
  message: false
  cache: true
---

```{r setup, include = FALSE}


# Set global chunk options
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 6,
  fig.height = 4,
  dpi = 300
)
```

```{r libraries}

# Load required R libraries
library(tidyverse)    # Data manipulation and visualization
library(here)         # File path management
library(knitr)        # Document generation
library(kableExtra)   # Enhanced table formatting
library(ggplot2)      # Advanced plotting
library(dplyr)        # Data manipulation
library(readr)        # Data import
library(scales)       # Scale functions for visualization
library(patchwork)    # Combining plots
library(flextable)    # Professional tables for Word output



```

```{r constants}

# Define any custom functions
format_p_value <- function(p) {
  case_when(
    p < 0.001 ~ "< 0.001",
    p < 0.01 ~ sprintf("= %.3f", p),
    TRUE ~ sprintf("= %.2f", p)
  )
}
```

```{r load-data}
source("Global.R")

spp.names.use <- spp.names %>% dplyr::select()

```


```{r table-creation}
# State mortality
s.m.1 <- 



```




# Introduction

Your introduction should provide background context, review relevant literature, identify gaps in current knowledge, and clearly state your research objectives and hypotheses.

## Background and Rationale

Provide the scientific background and rationale for your study. Cite relevant literature using the format [@author2023] for in-text citations.

## Study Objectives

Clearly state your primary and secondary objectives or research questions.

# Methods

## Study Design

Describe your study design, including the overall approach and methodology.

## Participants

Describe your study population, inclusion/exclusion criteria, and recruitment methods.

## Data Collection

Detail your data collection procedures, instruments used, and any quality control measures.

## Statistical Analysis

Describe your statistical analysis plan, including software used, significance levels, and specific tests performed.

# Results



Describe your main findings here, referencing 



```{r statistical-tests}
```

The analysis revealed a significant difference between groups , with a mean difference of  units.

# Discussion

## Main Findings

Summarize your key findings and their interpretation in the context of existing literature.

## Limitations

Acknowledge study limitations and their potential impact on your findings.

## Implications and Future Directions

Discuss the implications of your findings for theory, practice, and future research.

# Conclusion

Provide a concise summary of your main findings and their significance.

# Acknowledgments

Acknowledge funding sources, collaborators, and others who contributed to the work.

# References

::: {#refs}
:::

# Supplementary Material

```{r supplementary-table}
```

```{r supplementary-figure}

```
